Deep Learning

### 2015-01-13 deepreview-v2.tex
made v2

\documentclass[10pts]{article}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{natbib}
%\usepackage{float}

\addtolength{\textwidth}{2cm}
%\addtolength{\textwidth}{1cm}
\addtolength{\textheight}{2cm}
%\addtolength{\textheight}{1cm}
\addtolength{\topmargin}{-1cm}
\addtolength{\oddsidemargin}{-1cm}
\addtolength{\evensidemargin}{-1cm}

\parskip 0.1in
\parindent 0.0in
\setcounter{secnumdepth}{1}
\title{Deep Learning}

@@

Deep neural networks that learn many layers of feature detectors have
dramatically improved the state-of-the-art for speech recognition, object
recognition, object detection, predicting the activity of drug molecules,
and many other tasks. Deep learning discovers intricate structure in large
datasets by using the backpropagation algorithm to compute
error-derivatives for the weights that define each feature detector and
then repeatedly changing these weights in proportion to their derivatives
averaged over small batches of training examples.  Deep recurrent neural
networks, in which each layer also receives input from its own state at the
previous time-step, can now generate good captions for images and are
competitive for machine translation.

@@

Deep learning is currently making major advances in solving problems that
have resisted the best attempts of the Artificial Intelligence community
for many years. Deep learning is very good at discovering intricate
structure in large, high-dimensional datasets and is therefore widely
applicable in all of the sciences. For example, it has beaten previous
machine learning techniques at predicting the activity of potential drug
molecules~\citep{Dahl-et-al-arxiv2014}, detecting new fundamental particles~\citep{YANNparticles},
classifying galaxies~\citep{YANNgalaxies}, and predicting the effects of mutations in
non-coding DNA on gene expression and disease~\citep{Keung-et-al-2014}. % Frey

So far, most applications of deep learning use feed-forward neural
networks, but for tasks that involve sequential inputs it is often better
to use recurrent neural networks. These are given one element from an input % YB term-->element
sequence at each time step and their recurrent internal ``hidden'' state
depends on the whole history of the input sequence. Recurrent networks are
very good at predicting the next character in a string~\citep{Sutskever-et-al-ICML2011} or
the next word in a sequence~\citep{Mikolov-et-al-NIPS2013} but they can also be used for
more complex tasks.  After reading an English sentence one word at a time,
for example, an English``encoder'' network can be trained so that the final
state vector of its hidden neurons is a good representation of the thought
expressed by the sentence.  This thought vector can then be used as the
initial hidden state (or as extra input) of a trained French ``decoder'' 
network which outputs
a probability distribution for the first word of the French translation. If
a particular first word is chosen (e.g., according to that probability)
and provided as input to the decoder
network it will then output a probability distribution for the second word
of the translation and so on until a full stop is 
chosen~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}. 
Overall, this process generates sequences of French words according to
a probability distribution that depends on the English sentence.

@@

From the earliest days of pattern recognition~\citep{selfridge,Rosenblatt57}
researchers aimed to replace hand-engineered features by multiple layers of
learned features.  This was eventually achieved by using the
backpropagation algorithm which was discovered independently by several
different groups in the 1970s and 1980s~\citep{Werbos74,Parker85,LeCun85,RHW}.  
Backpropagation is simply an application of
the chain rule of derivatives (left of Figure~\ref{fig:backprop-box}),
which yields an efficient way of computing the gradient of a
smooth objective function with respect to all of the weights in a layered
network (Figures~\ref{fig:backprop-box} and~\ref{fig:mlp}). 
The objective function typically measures an error to be minimized: 
the discrepancy between
the actual output of the network on a given training case and the target
output for that case.  The key insight is that the error derivatives with
respect to the {\it outputs} of a layer can be computed by working
backwards from the error derivatives with respect to the outputs of the
subsequent layer. Once this backwards recursion has been used to compute
error derivatives for the outputs of all layers, it is easy to compute
error derivatives for the parameters (or weights) of each layer 
(Figure~\ref{fig:mlp}).

%%%%%%% GEOFF %%%%%%%%%%
%GEOFF: Do box explaining backprop equatons
%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htp]
\fbox{
\begin{minipage}{\textwidth}
\centerline{
\includegraphics[width=0.33\linewidth]{chain-rule.pdf}
\includegraphics[width=0.33\linewidth]{forward-graph.pdf}
\includegraphics[width=0.33\linewidth]{backward-graph.pdf}
}
\caption{
Left: the chain rule of derivatives tells us how
two small effects (that of a small change of $x$ on $y$, and that of $y$ on $z$)
are composed. A small change $\Delta x$ in $x$
gets transformed first into a small change $\Delta y$ in $y$
by getting multiplied by $\frac{\partial y}{\partial x}$
(that is the definition of partial derivative). Similarly,
the change $\Delta y$ creates a change $\Delta z$ in $z$. Substituting one equation
into the other gives the chain rule of derivatives, i.e., how
$\Delta x$ gets turned into $\Delta z$ through multiplication
by the product of $\frac{\partial y}{\partial x}$ and
$\frac{\partial z}{\partial x}$. It also works when $x$, $y$
and $z$ are vectors (and the derivatives are Jacobian matrices). \newline
Middle: many computations, such as the succession of linear and non-linear 
transformations of multi-layer neural networks, can be described as
a graph where each node $i$ computes a quantity $u_i$ (e.g. the
output of an artificial neuron or the value of a parameter such as
a synaptic weight), based on the set of values $a_i$, composed of elements $u_j$ of
predecessors in the graph, i.e., $u_i = f_i(a_i)$. This can be readily
generalized to nodes whose output is a vector or a matrix.\newline
Right: the chain rule can be applied efficiently to computational
flow graphs by recursively computing the derivative of the final
node $u_N$ (e.g., the error to be minimized or the output of the neural network)
with respect to every internal node. The basic recursive equation
is
$\frac{\partial u_N}{\partial u_j} = \sum_{i: j \in a_i} \frac{\partial u_N}{\partial u_i}
\frac{\partial f(a_i)}{\partial u_j}.$
}
\label{fig:backprop-box}
\end{minipage}
}
\end{figure}

@@

Backpropagation provides an efficient way of computing gradients with respect to all of
the weights in a layered network on a single training case, but there are
still many ways to make use of these gradients. The most obvious method is
to average the gradients over all of the cases in the training set and then
to change each weight in proportion to its negative gradient, which is the
direction where a small change will most improve the objective function. 
When the training set contains millions of cases and is highly redundant
% explaining what ``redundant'' means, here
(with many examples illustrating variations of the same underlying 
aspect or concept), however,
this is needlessly slow. It is much faster to compute the average gradient
on a small ``mini-batch'' of randomly chosen training cases and to update
each weight by this stochastic estimate of its full gradient, times a
learning rate which decays as learning proceeds. Stochastic gradient
descent can be improved by using the history of the individual weight
gradients to adapt the learning rate for each weight~\citep{bottou-bousquet-2008-small}
or to provide additional increments that depend on earlier gradients 
but these techniques are still the subject of controversy and are beyond
the scope of this review (but see~\citet{Montavon2012} for tricks of the trade). 
However, it is now widely accepted that some
form of stochastic gradient descent is the best bet for large, redundant
datasets.

@@

\section{Convolutional neural networks}

%%%%%%%% YANN %%%%%%%

YANN: can you summarize convnets and their use for reading cheques in about
350 words?

\citep{LeCun89-small}

@@

Recurrent neural networks are very powerful dynamical systems, but training
them with BPTT proved to be problematic, because the backpropagated
gradients either grow or shrink at each time step, so over many time-steps
they typically explode or vanish~\citep{Hochreiter91-small,Bengio-et-al-TNN1994},
and the conditions which allow stable dynamics also correspond to
vanishing gradients~\citep{Bengio-et-al-TNN1994},

\begin{figure}[ht]
\fbox{
\begin{minipage}{\textwidth}
\centerline{
\includegraphics[width=0.49\linewidth]{bp-mlp1.pdf}
\includegraphics[width=0.49\linewidth]{bp-mlp5.pdf}
}
\centerline{
\includegraphics[width=0.89\linewidth]{fig-hidden-recurrence-rnn.pdf}
}
\caption{Top left: graph of feedforward computations in a multi-layer
neural network with one input layer ($x$), two hidden layers ($h_1$ and $h_2$),
one output layer (out), a target $y$, and an error that compares the output
and the target. The synaptic weight matrices $W_1$, $W_2$ and $W_3$ are
the parameters of the neural network.\newline
Top right: back-propagation through the same multi-layer neural network
(downward arrows compute the partial derivative of the error with respect
to every node, i.e., the units value and the parameters).\newline
Bottom: a recurrent neural network (left) and the unfolding in time of the
computation involved in its forward computation. The artificial neurons
(e.g. hidden units grouped under node $s$ with values $s_t$ at time $t$) 
get inputs from other neurons at previous time steps
(this is represented with the black square, standing for a delay of
one time step, on the left). In this way, a recurrent net can map an input sequence
with elements $x_t$ into an output sequence with elements $o_t$, with
each $o_t$ depending on all the previous $x_{t'}$ (for $t'\leq t$). The
same parameters (matrices $U$, $V$, $W$) are used at each time step.
Many other architectures are possible, including a variant in which
the sampled outputs (e.g. words) can be generated according to an output 
distribution (the probability of the next word, for every possible next word) 
and used as inputs for the next time step.
The back-propagation algorithm (Figure~\ref{fig:backprop-box})
can be directly applied to the computational graph of the unfolded network
on the right hand side, to compute the derivative of a total error
(e.g., the log-probability of generating the right sequence of
words) with respect to all the states $s_t$ and all the parameters.
}
\label{fig:mlp}
\end{minipage}
}
\end{figure}

@@

In the 1990s, the machine learning community decided that backpropagation
in deep neural networks did not work well in practice and it was largely
abandoned.  At the time, people believed that learning many layers of
non-linear feature detectors with no prior knowledge was much too difficult
for a simple gradient descent procedure and that the networks were getting
trapped in poor local minima. We now know that the main reasons for the
limited success of backpropagation in deep networks were quite different:
The labeled datasets were far too small, the computers were far too slow,
and the weights were not initialised sensibly~\citep{GlorotAISTATS2010-small}.

The obvious weakness of gradient descent is that it can get trapped in poor
local minima, but in high-dimensional weight-spaces this is much less
common than getting stuck close to saddle points where the error surface
curves upwards in most directions but downwards in a few. Recent
results~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-arxiv2014} strongly suggest
the vast majority of critical points (where derivatives of the total
objective function with respect to parameters vanish) are saddle points, not
local minima, and that in large deep networks almost all of these local minima 
give very similar performance so it does not matter which one we find.  This goes a long way
towards explaining why stochastic gradient descent works so well for deep
neural networks.

@@

The first major application of this pre-training approach was in speech
recognition and it was made possible by the advent of fast Graphics
Processing Units (GPUs) that were convenient to
program~\citep{RainaICML09-small}.  In 2009, the approach was used to map
short temporal windows of coefficients extracted from a soundwave to a set
of probabilities for the various fragments of speech that might be
represented by the frame in the center of the window.  It achieved
record-breaking results on a standard speech recognition benchmark that
used a small vocabulary~\citep{TIMITpaper} and was quickly developed to
give record-breaking results on a large vocabulary task~\citep{Dahl2012}.
By 2012, versions of the deep net from 2009 were being developed by many of
the major speech groups~\citep{Hinton-et-al-2012} and were already being
deployed in Android phones.  For smaller datasets, unsupervised
pre-training helps to prevent overfitting~\citep{Erhan+al-2010-small},
leading to significantly better generalization when the number of labeled
examples is small, or in a transfer setting, i.e., when we have lots of
examples for some ``source'' tasks but very few for some ``target''
task. In this context, unsupervised representation learning helped to win
several transfer learning
competitions~\citep{UTLC+LISA-2011-small,Goodfellow-icml2012}.  Once deep
learning had been rehabilitated, it turned out that large labeled datasets
did not need the pre-training stage, provided the scales of the initial
random weights were carefully chosen to avoid very large or very small
initial gradients~\citep{GlorotAISTATS2010-small} and that even better
results (especially for very deep networks)
could be obtained by using rectifying non-linearities rather than
sigmoid or tanh
non-linearities~\citep{Glorot+al-AI-2011-small,Krizhevsky-2012-small}.

Interest in recurrent networks (which are deep in time) was revived by
``Long Short Term Memory'' networks (LSTMs) that use special hidden units
whose natural behaviour is to remember inputs for a long
time~\citep{Hochreiter+Schmidhuber-1997}.  
A linear ``memory cell'' interacts with other units
via gated connections whose weights are multiplied by the state of a
logistic gating unit that has an activation between $0$ and $1$. The input
vector has little effect on the state of that cell if the cell's input
gate has a low activation so memory cells can learn to selectively ignore
the input.  Similarly, a memory cell has little effect on the output if its
output gate has low activation. Each memory cell is a gated leaky neuron, i.e.,
it also has a connection to
itself at the next time-step that has a weight of 1 (so it tends to copy its
own real-valued state and accumulate the external signal), 
but this self-connection is {\em gated} by a ``forget gate'' so
the stored value leaks quickly unless the forget gate has an activation
near $1$.  The gates themselves have their activities determined by
weighted inputs that they receive from the input vectors and all of the
memory cells.

When LSTMs were first described % in English
\citep{Hochreiter+Schmidhuber-1997},
their complexity might have been off-putting
and, despite their impressive ability to learn very long-range temporal
structure on toy problems, they were largely ignored until they achieved
excellent results in a competition to read cursive
hand-writing~\citep{Graves-et-al-2009}. They have subsequently proved to be very
effective for tasks that have sequential input.  Combined with depth of
representation, they are now better than
deep feed-forward networks for speech recognition~\citep{Graves-et-al-ICASSP2013} and
unlike feed-forward nets they can implement an entire speech recognition
system that goes all the way from coefficents derived from the soundwave to
the sequence of characters in the transcription.  LSTMs or related forms of gated
units are also what is currently used for the encoder and decoder networks that
perform so well at machine
translation~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.

Recent work~\citep{Chung-et-al-NIPSDL2014-small,Yao-et-al-SLU-workshop2014} 
suggests that not all the elements of the LSTM
architecture are required, the crucial element being the presence and learnability 
of the forget gate, controlling the leaking rate (or time constant) of the cell.
%It is not yet clear which aspects of the complicated architecture of LSTMs
%are really necessary. Memory cells are a good way of allowing the
%backpropagated error derivatives to neither explode nor vanish, but they
%are not the only way.  For example, recurrent nets with rectified linear
%hidden units that use the non-linear function $y=max(0,x)$ are much simpler
%than LSTMs and if the recurrent weight matrix is initialized to be close to
%the identity matrix they work about as well as LSTMs both for toy problems
%involving very long-range temporal structure and for real tasks like
%predicting the next word in text\citep{LeHinton}.

\section{Object recognition and detection with deep convolutional networks}

When deep convolutional networks were applied to a dataset of about a
million images from the web that contained a thousand different classes
they achieved spectacular results, almost halving the error rates of the
best computer vision sytems~\citep{Krizhevsky-2012-small}.  
%Some of the images and the
%network's predictions are shown in figure~\ref{fig:imagenet-outputs}.  
The same kind of network is used to generate the captions shown in Figure~\ref{fig:caption-generation}.
The major factors in this
success were the large labeled dataset~\citep{imagenet_cvpr09}, the use of fast GPUs,
and the convolutional early layers.  It had already been shown that deep
supervised nets with rectified linear units could be trained without
unsupervised pre-training~\citep{Glorot+al-AI-2011-small}. 
These rectified linear units made the
training much faster and the fully connected final layers were prevented
from overfitting the peculiarities of the training set by randomly omitting
hidden units during training~\citep{Srivastava14}. Deep convolutional nets
have quickly replaced the carefully engineered pipelines full of
hand-designed features developed by the computer vision community and they
are now deployed in many computer vision applications, including image
tagging, object detection, face recognition, facial expression recognition
and scene parsing.

%%%%%%%% YANN %%%%%%%
YANN: Add about 300 words plus another part of the figure showing detected
parts of people.
%%%%%%%%%%%%%%%%%%%%

Cite at least \citep{LeCun+98} and \citep{Tompson-et-al-arxiv2014}

>>>>

\documentclass[]{article}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{float}
\parskip 0.1in
\parindent 0.0in
\setcounter{secnumdepth}{1}
\title{Deep Learning}

@@

Deep neural networks that learn to represent data in multiple layers
of increasing abstration have dramatically improved the
state-of-the-art for speech recognition, object recognition, object
detection, predicting the activity of drug molecules, and many other
tasks. Deep learning discovers intricate structure in large datasets
by using the backpropagation algorithm to compute error-derivatives
for the weights that define each layer of feature detector and then
repeatedly changing these weights in proportion to their derivatives
averaged over small batches of training examples.  Deep recurrent
neural networks, in which each layer also receives input from its own
state at the previous time-step, can now generate good captions for
images and are competitive for language translation.

@@

Deep learning is currently making major advances in solving problems
that have resisted the best attempts of the Artificial Intelligence
community for many years. Deep learning is very good at discovering
intricate structure in large, high-dimensional datasets and is
therefore applicable to many domains of science and business. For
example, it has beaten previous machine learning techniques at
predicting the activity of potential drug
molecules~\citep{DahlKaggle}, detecting new fundamental
particles~\citep{YANNparticles}, classifying
galaxies~\citep{YANNgalaxies}, and predicting the effects of mutations
in non-coding DNA on gene expression and
disease~\citep{Keung-et-al-2014}. % Frey

So far, most applications of deep learning use feed-forward neural
networks, but for tasks that involve sequential inputs, such as speech
and language, it is often better to use recurrent neural
networks. These are given one element from an input 
% YB term-->element 
sequence at each time step and their recurrent internal
``hidden'' state depends on the whole history of the input
sequence. Recurrent networks are very good at predicting the next
character in text~\citep{Sutskever-et-al-ICML2011} or the next
word in a sequence~\citep{Mikolov-et-al-NIPS2013} but they can also be
used for more complex tasks.  After reading an English sentence one
word at a time, for example, an English``encoder'' network can be
trained so that the final state vector of its hidden units is a good
representation of the thought expressed by the sentence.  This thought
vector can then be used as the initial hidden state (or as extra
input) of a trained French ``decoder'' network which outputs a
probability distribution for the first word of the French
translation. If a particular first word is chosen (e.g., according to
that probability) and provided as input to the decoder network it will
then output a probability distribution for the second word of the
translation and so on until a full stop is
chosen~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.
Overall, this process generates sequences of French words according to
a probability distribution that depends on the English sentence.

@@

From the earliest days of pattern recognition~\citep{selfridge,Rosenblatt57}
researchers aimed to replace hand-engineered features by multiple layers of
learned features.  This was eventually achieved by using the
backpropagation algorithm which was discovered independently by several
different groups in the 1970s and 1980s~\citep{Werbos74,Parker85,LeCun85,RHW}.  
Backpropagation is simply an application of
the chain rule of derivatives (left of Figure~\ref{fig:backprop-box}),
which yields an efficient way of computing the gradient of a
smooth objective function with respect to all of the weights and parameters in 
a layered network (Figures~\ref{fig:backprop-box} and~\ref{fig:mlp}). 
The objective function typically measures an error to be minimized: 
the discrepancy between
the actual output of the network on a given training case and the target
output for that case.  The key insight is that the error derivatives with
respect to the {\it inputs} of a layer can be computed by working
backwards from the error derivatives with respect to its {\em outputs}. 
Once this backwards recursion has been used to compute
error derivatives for the outputs of all layers, it is easy to compute
error derivatives for the parameters (or weights) of each layer 
(Figure~\ref{fig:mlp}).

%%%%%%% GEOFF %%%%%%%%%%
%GEOFF: Do box explaining backprop equatons
%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[H]
\fbox{
\begin{minipage}{\textwidth}
\centerline{
\includegraphics[width=0.33\linewidth]{chain-rule.pdf}
\includegraphics[width=0.33\linewidth]{forward-graph.pdf}
\includegraphics[width=0.33\linewidth]{backward-graph.pdf}
}
\caption{
Left: the chain rule of derivatives tells us how
two small effects (that of a small change of $x$ on $y$, and that of $y$ on $z$)
are composed. A small change $\Delta x$ in $x$
gets transformed first into a small change $\Delta y$ in $y$
by getting multiplied by $\frac{\partial y}{\partial x}$
(that is the definition of partial derivative). Similarly,
the change $\Delta y$ creates a change $\Delta z$ in $z$. Substituting one equation
into the other gives the chain rule of derivatives, i.e., how
$\Delta x$ gets turned into $\Delta z$ through multiplication
by the product of $\frac{\partial y}{\partial x}$ and
$\frac{\partial z}{\partial x}$. It also works when $x$, $y$
and $z$ are vectors (and the derivatives are Jacobian matrices). \newline
Middle: many computations, such as the succession of linear and non-linear 
transformations of multi-layer neural networks, can be described as
a graph where each node $i$ computes a quantity $u_i$ (e.g. the
output of an artificial neuron or the value of a parameter such as
a synaptic weight), based on the set of values $a_i$, composed of elements $u_j$ of
predecessors in the graph, i.e., $u_i = f_i(a_i)$. This can be readily
generalized to nodes whose output is a vector or a matrix.\newline
Right: the chain rule can be applied efficiently to computational
flow graphs by recursively computing the derivative of the final
node $u_N$ (e.g., the error to be minimized or the output of the neural network)
with respect to every internal node. The basic recursive equation
is
$\frac{\partial u_N}{\partial u_j} = \sum_{i: j \in a_i} \frac{\partial u_N}{\partial u_i}
\frac{\partial f(a_i)}{\partial u_j}.$
}
\end{minipage}
}
\label{fig:backprop-box}
\end{figure}

@@

Backpropagation provides an efficient way of computing gradients with respect to all of
the weights in a layered network on a single training case, but there are
still many ways to make use of these gradients. The most obvious method is
to average the gradients over all of the cases in the training set and then
to change each weight in proportion to its negative gradient, which is the
direction where a small change will most decrease the objective function. 
When the training set contains millions of cases and is highly redundant
% explaining what ``redundant'' means, here
(with many examples illustrating variations of the same underlying 
aspect or concept), however,
this is needlessly slow. It is much faster to compute the average gradient
on a small ``mini-batch'' of randomly chosen training cases and to update
each weight by this stochastic estimate of its full gradient, times a
learning rate which decays as learning proceeds. Stochastic gradient
descent can be improved by using the history of the individual weight
gradients to adapt the learning rate for each weight~\citep{bottou-bousquet-2008-small}
or to provide additional increments that depend on earlier gradients 
but these techniques are still the subject of controversy and are beyond
the scope of this review (but see~\citet{Montavon2012} for tricks of the trade). 
However, it is now widely accepted that some
form of stochastic gradient descent is the best bet for large, redundant
datasets.

@@

\section{Convolutional networks}

%%%%%%%% YANN %%%%%%% 
% YANN: can you summarize convnets and their use for reading cheques in about
% 350 words?
% I couldn't do it in 350 words.

\vspace{-0.12cm}
\subsection*{Convolutional Networks}
\vspace{-0.12cm}

\begin{figure}
\begin{center}
  \vspace{0.2cm}
  % \includegraphics[width=0.99\linewidth]{convnet-diagram}
  nice convnet figure goes here
\end{center}
\vspace{-0.6cm}
\caption{A typical ConvNet architecture}
\vspace{-0.8cm}
\label{fig:convnet}
\end{figure}

One of the most widely used deep learning architectures is the
convolutional networt, or ConvNet~\cite{lecun-90c,lecun-98}. ConvNets
are designed to process data that come in the form of arrays, for
example a color image composed of three 2D arrays containing pixel
intensities in the three color channels. Many data modalities are in
the form of multiple arrays: 1D for signals and language, 2D for
images or audio spectrograms, 3D for video or volumetric images.
There are four key ideas behind ConvNets that take advantage of the
properties of many such natural signals: local connections, shared
weights, pooling, and the use of many layers.

The architecture of a typical ConvNet is shown in
figure~\ref{fig-convnet}. The first few stages are composed of two
types of layers: convolutional layers and pooling layers. Units in a
convolutional layer are organized in {\em faeture maps}, within which
each unit is connected to local patches in the feature maps of the
previous layer through a set of weights called a {\em filter
  bank}. The result of this local weight sum is then passed through a
non-linearity.  All units in a feature map share the same filter
bank. Multiple feature maps in a layer use different filter banks. The
reason for this architecture is twofold. First, In array data such as
images, local groups of values are often highly correlated forming
local motifs that are easily detected. Second, the local statistics of
images and other signals are invariant to location. In other word, if
a motif can appear in one part of the image, it could appear anywhere,
hence the idea of units at different locations sharing the same
weights.  Mathematically, the filtering operation performed by a
feature map is a {\em discrete convolution}, hence the name
ConvNet. While the role of the convolutional layer is to detect local
conjunctions of features in the previous layer, the role of the
pooling layer is to merge semantically similar features into
one. Since the relative positions of the features forming a motif can
vary somewhat, reliably detecting the motif can be done by
coarse-graining the position of each feature. A typical pooling unit
computes the max of a local patch of units in one feature map (or in a
few). Neighboring pooling units take input from patches that are
shifted by more than one row or column, thereby reducing the dimension
of the representation and creating an invariance to small shifts.  Two
or three stages of convolutionas, non-linearity, and pooling are
stacked, followed by more convolutional and fully-connected layers.
Back-propagating gradients through a ConvNet is as simple
as through a regular deep network, allowing all the weights in all the
filter banks to be trained.

This architecture exploits the property that many natural signals are
{\em compositional hierarchies}. In images, local combinations of
edges form motifs, motifs assemble into parts, and parts form objects.
Similar hierarchies exists in speech and text from sounds to phones,
phonemes, syllables, words, and sentences. The elements that form
elements in the next level can vary in relative position and
appearance. The convolutional and pooling layers in ConvNets are
directly inspired by the classic notions of simple cells and complex
cells in visual neuroscience~\cite{Hubel1962}, and the overall
architecture is reminiscent of visual cortex ventral pathway hierarchy
LGN-V1-V2-V4-IT~\cite{VanEssen}. ConvNets have their roots in
Fukushima's Neocognitron~\cite{fukushima-82} whose architecture was
somewhat similar but did not have an end-to-end supervised learning
algorithm such as backpropagation+SGD.

There have been numerous applications of convolutional networks going
back to the early 1990s. Perhaps the earliest success was a bank check
reading system which was commercially deployed by AT\&T and NCR in
1994~\cite{lecun-98}. The system used a ConvNets trained jointly with
a graphical model that implemented language constraints. By the late
1990s this system was reading over 10\% of all the checks in the
US. Microsoft later deploy ConvNets in a number of OCR and handwriting
recognition
systems~\cite{simard-03,chellapilla-ist-06,chellapilla-iwfhr-06b}
including for Arabic and
Chinese~\cite{abdulkader-iwfhr-06,chellapilla-iwfhr-06a}. ConvNets
have also been used since the early 1990s for object detection in
natural images, including faces and
pedestrians~\cite{vaillant-monrocq-lecun-94,garcia-delakis-04,osadchy-07,nasse-09,sermanet-cvpr-13}.

@@

Recurrent neural networks are very powerful dynamical systems, but training
them with BPTT proved to be problematic, because the backpropagated
gradients either grow or shrink at each time step, so over many time-steps
they typically explode or vanish~\citep{Bengio-et-al-TNN1994,Hochreiter91-small}.

\begin{figure}[H]
\fbox{
\begin{minipage}{\textwidth}
\centerline{
\includegraphics[width=0.49\linewidth]{bp-mlp1.pdf}
\includegraphics[width=0.49\linewidth]{bp-mlp5.pdf}
}
\centerline{
\includegraphics[width=0.89\linewidth]{fig-hidden-recurrence-rnn.pdf}
}
\caption{Top left: graph of feedforward computations in a multi-layer
neural network with one input layer ($x$), two hidden layers ($h_1$ and $h_2$),
one output layer (out), a target $y$, and an error that compares the output
and the target. The synaptic weight matrices $W_1$, $W_2$ and $W_3$ are
the parameters of the neural network.\newline
Top right: back-propagation through the same multi-layer neural network
(downward arrows compute the partial derivative of the error with respect
to every node, i.e., the units value and the parameters).\newline
Bottom: a recurrent neural network (left) and the unfolding in time of the
computation involved in its forward computation. The artificial neurons
(e.g. hidden units grouped under node $s$ with values $s_t$ at time $t$) 
get inputs from other neurons at previous time steps
(this is represented with the black square, for delay of
one time step, on the left). In this way, a recurrent net can map an input sequence
with elements $x_t$ into an output sequence with elements $o_t$, with
each $o_t$ depending on all the previous $x_{t'}$ (for $t'\leq t$). The
same parameters (matrices $U$, $V$, $W$) are used at each time step.
Many other architectures are possible, including a variant in which
the outputs (e.g. words) can be generated according to a distribution 
(the probability of the next word) and used as inputs for the next time step.
The back-propagation algorithm (Figure~\ref{fig:backprop-box})
can be directly applied to the computational graph of the unfolded network
on the right hand side, to compute the derivative of a total error
(at all time steps) with respect to all the states $s_t$ and all the parameters.
}
\end{minipage}
}
\label{fig:mlp}
\end{figure}

@@

In the 1990s, the machine learning community decided that backpropagation
in deep neural networks did not work well in practice and it was largely
abandoned.  At the time, people believed that learning many layers of
non-linear feature detectors with no prior knowledge was much too difficult
for a simple gradient descent procedure and that the networks were getting
trapped in poor local minima. We now know that the main reasons for the
limited success of backpropagation in deep networks were quite different:
The labeled datasets were far too small, the computers were far too slow,
and the weights were not initialised sensibly.

The obvious weakness of gradient descent is that it can get trapped in poor
local minima, but in high-dimensional weight-spaces this is much less
common than getting stuck close to saddle points where the error surface
curves upwards in most directions but downwards in a few. Recent
results~\citep{Dauphin-et-al-NIPS2014-small,Choromanska-et-al-arxiv2014} strongly suggest
that in deep networks almost all of these local minima give very similar
performance so it does not matter which one we find.  This goes a long way
towards explaining why stochastic gradient descent works so well for deep
neural networks.

@@

The first major application of this pre-training approach was in speech
recognition and it was made possible by the advent of fast Graphics
Processing Units (GPUs) that were convenient to program~\citep{RainaICML09-small}.  
In 2009, the approach was used to map short temporal windows of coefficients extracted
from a soundwave to a set of probabilities for the various fragments of
speech that might be represented by the frame in the center of the window.
It achieved record-breaking results on a standard speech recognition
benchmark that used a small vocabulary~\citep{TIMITpaper} and was quickly
developed to give record-breaking results on a large vocabulary
task~\citep{Dahl}.  By 2012, versions of the deep net from 2009 were being
developed by many of the major speech groups\citep{Hinton-et-al-2012} and were already
being deployed in Android phones.  For smaller datasets, unsupervised
pre-training helps to prevent overfitting~\citep{Erhan+al-2010-small}, 
leading to significantly better generalization when the number of
labeled examples is small. Once deep
learning had been rehabilitated, it turned out that large labeled datasets
did not need the pre-training stage, provided the scales of the initial
random weights were carefully chosen to avoid very large or very small
initial gradients.

Interest in recurrent networks (which are deep in time) was revived by
``Long Short Term Memory'' networks (LSTMs) that use special hidden units
whose natural behaviour is to remember inputs for a long
time~\citep{Hochreiter+Schmidhuber-1997}.  
A linear ``memory cell'' interacts with other units
via gated connections whose weights are multiplied by the state of a
logistic gating unit that has an activation between $0$ and $1$. The input
vector has little effect on the state of that cell if the cell's input
gate has a low activation so memory cells can learn to selectively ignore
the input.  Similarly, a memory cell has little effect on the output if its
output gate has low activation. Each memory cell also has a connection to
itself (it is a leaky neuron) at the next time-step that has a weight of 1 so it tends to copy its
own real-valued state, but this self-connection is gated by a ``keep gate'' so
the stored value leaks quickly unless the keep-gate has an activation
near $1$.  The gates themselves have their activities determined by
weighted inputs that they receive from the input vectors and all of the
memory cells.

When LSTMs were first described % in English
\citep{Hochreiter+Schmidhuber-1997} their complexity was off-putting
and, despite their impressive ability to learn very long-range temporal
structure on toy problems, they were largely ignored until they achieved
excellent results in a competition to read cursive
hand-writing~\citep{Graves-et-al-2009}. They have subsequently proved to be very
effective for tasks that have sequential input.  They are now better than
deep feed-forward networks for speech recognition~\citep{Graves-et-al-ICASSP2013} and
unlike feed-forward nets they can implement an entire speech recognition
system that goes all the way from coefficents derived from the soundwave to
the sequence of characters in the transcription.  They
are also what is currently used for the encoder and decoder networks that
perform so well at machine
translation~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.

Recent work~\citep{Chung-et-al-NIPSDL2014-small,Yao-et-al-SLU-workshop2014} 
suggests that not all the elements of the LSTM
architecture are required, the crucial element being the presence and learnability 
of the so-called forget gate, controlling the leaking rate (or time constant) of the cell.
%It is not yet clear which aspects of the complicated architecture of LSTMs
%are really necessary. Memory cells are a good way of allowing the
%backpropagated error derivatives to neither explode nor vanish, but they
%are not the only way.  For example, recurrent nets with rectified linear
%hidden units that use the non-linear function $y=max(0,x)$ are much simpler
%than LSTMs and if the recurrent weight matrix is initialized to be close to
%the identity matrix they work about as well as LSTMs both for toy problems
%involving very long-range temporal structure and for real tasks like
%predicting the next word in text\citep{LeHinton}.

\section{Image understanding with deep convolutional networks}

Since the early 2000s, there have been many tasks in which ConvNets
have been applied with success to the detection, segmentation, and
recognition of objects and regions in images. These were invariably
tasks in which labeled data was relatively abundant, such as traffic
sign recognition~\cite{sermanet-ijcnn-11,schmidhuber-traffic}, the
segmentation of biological images~\cite{ning-05} particularly for
connectomics~\cite{seung-xx}, the detection of faces, text, and human bodies in natural
images~\cite{vaillant-monrocq-lecun-94,nowlan-platt-95,garcia-delakis-04,osadchy-07,nasse-09,taylor-xx,sermanet-cvpr-13,thompson-14}.

An important application is the labeling of images at the pixel level,
whose applications include driving mobile robots and self-driving
cars~\cite{hadsell-jfr-09,farabet-icml-12},

Other applications of ConvNets that are gaining importance concern
natural language understanding~\cite{collobert-weston-xx}, and speech
recognition~\cite{sainath-xx}.

But works were largely ignored by the computer vision and machine
learning community until the ImageNet competition in 2012.

When deep convolutional networks were applied to a dataset of about a
million images from the web that contained a thousand different classes
they achieved spectacular results, almost halving the error rates of the
best computer vision sytems~\citep{Krizhevsky-2012-small}.  Some of the images and the
network's predictions are shown in figure~\ref{fig:imagenet-outputs}.  
The major factors in this
success were the large labeled dataset~\citep{imagenet_cvpr09}, the use of fast GPUs,
and the convolutional early layers.  It had already been shown that deep
supervised nets with rectified linear units could be trained without
unsupervised pre-training~\citep{Glorot+al-AI-2011-small}. 
These rectified linear units made the
training much faster and the fully connected final layers were prevented
from overfitting the peculiarities of the training set by randomly omitting
hidden units during training~\citep{Srivastava14}. Deep convolutional nets
have quickly replaced the carefully engineered pipelines full of
hand-designed features developed by the computer vision community and they
are now deployed in many computer vision applications, including image
tagging, object detection, face recognition, facial expression recognition
and scene parsing.
